Clustering Constraints on the Relative Sizes of Central and Satellite Galaxies

### 2017-11-23 paper/galsize_paper.tex
AK's edits

\author[Hearin, Behroozi, Kravtsov \& Moster]{
Andrew Hearin$^{1}$, Peter Behroozi$^{2}$, Andrey Kravtsov$^{3}$, Benjamin Moster$^{4,5}$\\
$^{1}$Argonne National Laboratory, Argonne, IL, USA 60439, USA\\
$^{2}$Department of Physics, University of Arizona, 1118 E 4th St, Tucson, AZ 85721 USA\\
$^{3}$Department of Astronomy \& Astrophysics, and Kavli Institute for Cosmological Physics, University of Chicago, Chicago IL 60637\\
$^{4}$Universit{\"a}ts-Sternwarte, Ludwig-Maximilians-Universit{\"a}t M{\"u}nchen, Scheinerstr. 1, 81679 M{\"u}nchen, Germany\\
$^5$Max-Planck Institut f\"ur Astrophysik, Karl-Schwarzschild Stra\ss e 1, 85748 Garching, Germany
}

@@

\begin{abstract}
We place empirical constraints on the connection between dark matter halos and galaxy half-light radii, $\rhalf.$ Low-redshift SDSS measurements show that smaller galaxies cluster much more strongly than larger galaxies at fixed stellar mass. Using {\tt Halotools} to forward model the observations, we find that the clustering signal generically requires satellite galaxies to be smaller than central galaxies of the same halo mass. We present a simple empirical model consistent with the clustering results, in which galaxy size is proportional to halo virial radius at the time of peak halo mass. We use this model to predict how galaxy lensing, $\Delta\Sigma,$ should depend on $\rhalf$ for $\mstar-$complete samples. Other simple empirical models fail the clustering test, such as models in which galaxy size is related to stellar mass alone; these failures persist even when accounting for possible effects from satellite stripping and orphan galaxies. Our results suggest that the relative size of centrals and satellites could be predetermined at the time of satellite infall, and that a remarkably simple galaxy--halo scaling relation emerges from the complex physics regulating galaxy size.
\end{abstract}

@@

Observational constraints on models for galaxy $\rhalf$ typically come from one-point function measurements such as $\mean{\rhalf}{\mstar}$ and $\sigma(\rhalf\vert\mstar),$ e.g.,  \citet{khochfar_silk06,lang_etal14,desmond_etal17,bottrell_etal17b,hou_etal17,somerville_etal17}, or otherwise from catalogs of galaxy groups \citep{weinmann_etal08,guo_etal09,huertas_company_etal13b,spindler_wake17}. More recently, the connection between a galaxy's size and the virial radius of its parent halo has been
explored by converting stellar mass to halo mass and virial radius, using abundance matching \citep{kravtsov_etal04,tasitsiomi_etal04,vale_ostriker04,vale_ostriker06,conroy_etal06}.  Such analyses have indicated that galaxy size is, on average, proportional to host halo virial radius, both at $z=0$ \citep{kravtsov13}, and higher redshifts \citep{huang_etal17,somerville_etal17}.

@@

The approach we take is to forward model the $\rhalf-$halo connection: we generate Monte Carlo realizations of models using a high-resolution cosmological N-body simulation, and then make synthetic measurements in a manner that closely mimics the measurements we make on our observed galaxy samples. We describe the dataset and simulations we use in \S\ref{sec:data}. Our forward-modeling methods are described in \S\ref{sec:model}, with supplementary material in the Appendix. All of our primary results appear in \S\ref{sec:results}, and their theoretical interpretation in \S\ref{sec:interpretation}. We contextualize our findings in terms of previous work in \S\ref{sec:previous_work}, and discuss future directions in \S\ref{sec:future}. We conclude the paper with a summary of our principal results in \S\ref{sec:conclusion}.

@@

As the bedrock of our modeling, we use the publicly available\footnote{\url{http://www.slac.stanford.edu/~behroozi/BPlanck\_Hlists}} catalog of {\tt Rockstar} subhalos identified at $z=0$ in the Bolshoi-Planck simulation \citep{klypin_etal11,klypin_etal16,behroozi12_rockstar,riebe_etal13,behroozi_etal12b,rodriguez_puebla16_bolplanck}. As described in \S\ref{subsec:sham}, we use traditional abundance matching to connect stellar mass $\mstar$ with subhalo mass. To address issues related to subhalo incompleteness \citep{guo_white13,campbell_etal17}, we supplement the Bolshoi-Planck subhalo catalog with subhalos that have been disrupted and no longer appear in the standard catalog as $z=0$ surviving subhalos. We describe our treatment of these ``orphan" galaxies in the Appendix, where we also show results that exclude all orphan galaxies.

For our SDSS galaxy samples, we calculate two-point clustering $\wproj$ using line-of-sight projection of $\pi_{\rm max}=20~\mpc$ using the {\tt correl} program in {\tt UniverseMachine} (Behroozi et al. 2017, in prep). The {\tt correl} code uses the \citet{landy_szalay93} estimator ($DD - 2DR + RR$) to compute the redshift-space correlation function $\xi(\rproj, \pi)$, which is then integrated over $\vert\pi\vert < 20$ Mpc to compute the projected correlation function $\wproj(\rproj)$.  The code uses $10^6$ randoms drawn from the same mask region with uniform volume distribution to compute $DR$, and $RR$ is computed via Monte Carlo integration. Errors are estimated by jackknife resampling, giving us an estimate for the lower bound of samples with volumes $V_{\rm eff}\lesssim 0.3{\rm Gpc}^3.$

@@

All numerical values of $\rhalf$ will be quoted in physical $\kpc,$ and all values of $\mstar$ and $\mhalo$ in $\msun,$ assuming $H_0=67.8~\kms\equiv100h~\kms,$ the best-fit value from \citet{planck15}. To scale stellar masses to ``$h=1$ units" \citep{croton13}, our numerically quoted values for $\mstar$ should be multiplied by a factor of $h^2,$ while our halo masses and distances should be multiplied by a factor of $h.$

@@

Because galaxy clustering has well-known dependence upon $\mstar$ that is not the subject of this work, we wish to remove this influence and focus purely on the relationship between $\rhalf$ and $\wproj(\rproj).$ To do so, we determine the value $\median{\rhalf}{\mstar}$ by computing a sliding median of $\rhalf$ for galaxies sorted by $\mstar,$ calculated using a window of width $N_{\rm gal}=1000.$ Each galaxy is categorized as either ``large" or ``small" according to whether it is above or below the median value appropriate for its stellar mass. We note that this is analogous to the common convention for studying the properties of ``red" vs. ``blue" galaxies, in which the two subsamples are divided by a $\mstar-$ or luminosity-dependent green valley cut \citep[e.g.,][]{vdB_etal08,zehavi_etal11}, only here the $\rhalf$ distribution is uni-modal, not bi-modal.

@@

We map $\mstar$ onto subhalos using deconvolution abundance matching (SHAM). See Section 3.3 of \citet{behroozi_etal10}, or the appendix of \citet{kravtsov_etal14}, for an account of the mathematics underlying this technique in the presence of scatter. To perform the abundance matching, we use the publicly available code\footnote{\url{https://bitbucket.org/yymao/abundancematching/overview}} developed by Yao-Yuan Mao \citep{lehmann_etal15} that provides a python wrapper around the C kernel developed in \citet{behroozi_etal10}. For the subhalo property used in SHAM, we use $\mpeak,$ the largest value of $\mvir$ ever attained by the subhalo throughout the entire history of the main progenitor halo.  Using the stellar mass function provided in \citet{moustakas_etal13}, we model $\mstar$ as a log-normal distribution with $0.2$dex of scatter about the median relation $\median{\mstar}{\mpeak}$ determined by SHAM \citep{reddick_etal13}. As described in the appendix, we include a prescription for supplementing the ordinary {\tt Rockstar} subhalo catalog with disrupted subhalos, so that some of our model galaxies inhabit subhalos that are no longer resolved and do not appear in the standard publicly available catalog (``orphan galaxies").

@@

\subsubsection{Tidal stripping model}
\label{subsubsec:strippingmodel}

In the first model we explore, we assume that stellar mass $\mstar$ controls $\rhalf.$ To implement this model, for simplicity we tabulate $\median{\rhalf}{\mstar}$ directly from the data, rather than pursue a parametric form, so that this model reproduces the observed $\rhalf-\mstar$ relation by construction. In order to ensure that the model has a level of scatter in $\rhalf$ at fixed $\mstar$ that is comparable to the level reported in \citet{somerville_etal17}, our model sizes are drawn from a log-normal distribution with $0.2$ dex of scatter centered at the observed $\median{\rhalf}{\mstar}.$

It is natural to consider the possibility that stellar mass is stripped from satellite galaxies after infall. The basis for our implementation of this phenomenon is the fitting function presented in \citet{smith_etal16}, which was calibrated by studying stellar mass loss in a suite of high-resolution hydrodynamical simulations. In this model, $f_{\ast}$ quantifies the remaining fraction of stellar mass as a function of $f_{\rm DM},$ the fraction of dark matter mass that remains after subhalo infall:
\beq
f_{\ast} = 1 - \exp(-14.2f_{\rm DM}).
\eeq
For $f_{\rm DM}$ we use the ratio of present-day subhalo mass divided by the peak mass, $M_{\rm vir}/M_{\rm peak}.$ If we denote the post-stripping stellar mass as $M_{\ast}',$ then we have $M_{\ast}'\equiv f_{\ast}M_{\ast},$ where $M_{\ast}$ is given by the initial application of abundance matching. We then calculate the post-stripping radius by interpolating $\langle\rhalf'\vert\mstar'\rangle$ directly from SDSS data. In this way, we model satellite mass loss and  size truncation in a manner that mimics what is seen in hydrodynamical simulations. As shown in \citet{smith_etal16}, stellar stripping in this model does not begin until $f_{\rm DM}\lesssim0.2,$ and in order for $\sim50\%$ of stars to be stripped, the DM halo needs to be stripped to $\sim5\%$ of its initial mass.

@@

As noted in \S\ref{sec:intro}, \citet{kravtsov13} showed that galaxy sizes are on average proportional to the virial radius of their host halo. Motivated by these results, we  explore a model in which $\rhalf$ scales linearly with halo virial radius:\footnote{We note that the normalization factor in \citet{kravtsov13} is 0.015, where as our normalization is 0.01. The difference derives from two factors: {\em i.} the choice of halo radius definition: \citet{bryan_norman98} in the present work, compared to the $200c$ definition used in \citet{kravtsov13}; {\em ii.} the latter study used deprojected (larger) sizes for spheroidal galaxies, whereas we make no such correction here. See \S\ref{sec:future} for further discussion.}
\beq
\label{eq:fiducial_model}
\rhalf = 0.01\rvir
\eeq
For the virial radius of halos and subhalos, we use $\rmpeak:$ the value of $\rvir$ in physical units of $\kpc$ measured at the time when halo mass reached its maximum, $\zpeak.$  The relationship between $\rmpeak, \mpeak$ and $\zpeak$ is given by
\beq
\mpeak\equiv\frac{4\pi}{3}\rmpeak^{3}\Delta_{\rm vir}(\zpeak)\rho_{\rm m}(\zpeak),
\eeq
where for $\Delta_{\rm vir}(\zpeak)$ we use the fitting function to the ``virial" definition used in \citet{bryan_norman98}. This modeling assumption effectively defines $\median{\rhalf}{\mhalo};$ this relation defines the center of the log-normal distribution with $0.2$ dex of scatter that we use to generate a Monto Carlo realization of the model population.

@@

The points and curves in Figure \ref{fig:clustering_ratio_upshot} are all negative: small galaxies cluster more strongly relative to large. This presentation of the measurement makes it clear that the underlying signal of $\rhalf-$dependent clustering is strongest for samples with smaller stellar mass; as $\mstar$ increases, the signal weakens and nearly vanishes for $\mstar\gtrsim10^{11}\msun.$ Strikingly, the $\rvir-$based model reproduces this $\mstar-$dependent behavior, as well as the scale-dependence of the observed clustering signal at each $\mstar.$ As described in \S\ref{subsec:censat_sizes} below, we attribute the success of this prediction to the systematic difference of sizes of central and satellite galaxies.

@@

Due to the shape of ${\rm d}\rhalf/{\rm d}\mstar,$ low-mass satellites have to lose much more halo mass than high-mass satellites in order to get pushed into the small-$\rhalf$ sample. Thus even though there are more satellites at lower masses, the shape of ${\rm d}\rhalf/{\rm d}\mstar$ counteracts this trend, making the mass-dependence of the observed signal shown in Figure \ref{fig:mstarmodelclustering} difficult to explain by appealing to tidal stripping. Evidently it is difficult to strip enough mass from satellites so that the $\rhalf-$dependent clustering predictions are in agreement with observations.  We reach the same conclusion when we tested an even more extreme $\mstar-$stripping model, in which stellar mass loss is linearly proportional to halo mass loss \citep[][Model 1]{watson_etal12}.
As we argue in \S\ref{sec:interpretation}, this supports the conclusion that the relative difference between central and satellite size is largely in place at the time of satellite infall.

@@

%---------------------------------------------------------------------------------------------------
\begin{figure}
\centering
\includegraphics[width=8cm]{FIGS/color_selected_size_dependent_clustering.pdf}
\caption{
{\bf Distinct $\rhalf-$dependence of clustering for color-selected galaxy samples.}
Projected clustering of SDSS galaxies in the bin of stellar mass: $10^{9.75}<\mstar/\msun<10^{10.25}.$ For both ``blue" and ``red" samples, we first divide SDSS galaxies falling in the $\mstar$ bin based on a color cut at $g-r=0.65.$ We then subsequently compute $\median{\rhalf}{\mstar; {\rm red}}$ and $\median{\rhalf}{\mstar; {\rm blue}}$. In analogy to the bottom left panel of Figure \ref{fig:sizedefinition}, for each of the red and blue samples, the small and large subsamples have statistically indistinguishable stellar mass functions. We then measure $\wproj$ for each of the four subsamples, plotting the results with color-coding as specified in the legend. At this stellar mass, pre-selecting galaxy samples by color results in a dramatic weakening of the $\rhalf-$dependence of $\wproj$ relative to $\mstar-$complete samples, as well as a sign-reversal of the signal. In \S\ref{sec:interpretation} we give a simple halo model picture unifying the trends shown here with those illustrated in Figure \ref{fig:clustering_ratio_upshot}.
}
\label{fig:colorclustering}
\end{figure}
%---------------------------------------------------------------------------------------------------

@@

Likewise, the significantly weaker dependence of clustering on size for color-selected samples (Figure \ref{fig:colorclustering}), and its reversed sign, can be naturally explained by the same principles.  It is well known that satellites are redder than centrals of the same stellar mass \citep[e.g.,][]{vdB_etal08}, and so blue galaxies have a small satellite fraction, while red galaxies have a large fraction of satellites that occupy high-mass halos; accordingly, red galaxies cluster much more strongly than blue galaxies \citep[e.g.,][]{zehavi_etal11}. Pre-selecting galaxy samples by color then naturally restricts the impact of satellites on size-dependent clustering. For blue galaxies, the reason traces to the small value of $F_{\rm sat};$ for red galaxies, the reason traces to the limited dynamic range of host halo masses occupied by the satellites. With the impact of satellites reduced, the relative clustering of centrals becomes more important. In the limit that size-dependent clustering is dominated by centrals, the $\rvir$-based size model predicts that large galaxy samples should be (weakly) more clustered than small ones, because $\rhalf\propto\rvir\propto\mhalo^{1/3}.$ This is the trend seen in the clustering measurements shown in Figure \ref{fig:colorclustering}.

@@

In the $\rvir$-based size model, the weak dependence of $\rhalf$ on $\mstar$ for low-mass galaxies is due to the steep $\mstar-\mvir$ relation. For example, if $\mstar\propto\mvir^\alpha\propto\rvir^{3\alpha}$, then $\rhalf\propto \rvir\propto \mstar^{1/(3\alpha)}$. Given that $\alpha\sim 1-3$ is inferred for the $\mstar-\mvir$ relation at low masses \citep{kravtsov10,moster_etal13,behroozi13_smhm,kravtsov_etal14}, the $\rhalf-\mstar$ relation in this regime is expected to be shallow.
For galaxies of $\mstar\gtrsim 5\times 10^{10}\ M_\odot$, $\alpha\approx 0.3$ \citep[e.g.,][]{kravtsov_etal14}, and thus
the slope of the $\rhalf-\mstar$ relation is approximately linear. These trends of $\rhalf$ with $\mstar$ have a direct relation to the trends of surface brightness $\mu_\star=\mstar/(\pi\rhalf^2)$ with stellar mass \citep[see, for example, the $\mu_\star-$dependent clustering measurements in][which are qualitatively similar to the measurements reported here]{li_etal06}.

@@

The right panels of Figure \ref{fig:lensingprediction} compare the distributions of host halo mass, $\mhost,$ for small and large samples;\footnote{We calculated the PDFs shown in Figure \ref{fig:lensingprediction} using the Scipy implementation of Gaussian kernel density estimation \citep{scipy}.} for centrals, $\mhost=\mhalo,$ while for satellites, $\mhost$ is the virial mass of the parent halo. The peak of each distribution arises from central galaxies, while the long tail at large values of $\mhost$ is due to satellites. At all stellar mass, the blue histograms peak at slightly larger $\mhost$ relative to the red; this is because the $\rvir-$based model predicts that $\mhalo$ increases with $\rhalf$ for centrals. For $\mstar-$complete samples, this effect only becomes dominant when the satellite fraction drops to sufficiently small values. Qualitatively, these features are completely generic predictions of the $\rvir-$based model; quantitatively, the precise value of $\mstar$ where the effect of centrals begins to dominate $\Delta\Sigma$ depends sensitively on $F_{\rm sat}(\mstar),$ which we have only roughly calibrated in the present work.

>>>>

\author[Hearin, Behroozi, Kravtsov \& Moster]{
Andrew Hearin$^{1}$, Peter Behroozi$^{2}$, Andrey Kravtsov$^{3}$, Benjamin Moster$^{4,5}$\\
$^{1}$Argonne National Laboratory, Argonne, IL, USA 60439, USA\\
$^{2}$Department of Physics, University of Arizona, 1118 E 4th St, Tucson, AZ 85721 USA\\
$^{3}$Department of Astronomy \& Astrophysics, and Kavli Institute for Cosmological Physics, The University of Chicago, Chicago IL 60637\\
$^{4}$Universit{\"a}ts-Sternwarte, Ludwig-Maximilians-Universit{\"a}t M{\"u}nchen, Scheinerstr. 1, 81679 M{\"u}nchen, Germany\\
$^5$Max-Planck Institut f\"ur Astrophysik, Karl-Schwarzschild Stra\ss e 1, 85748 Garching, Germany
}

@@

\begin{abstract}
We explore size dependence of galaxy clustering and the implied connection between galaxy half-light radii, $\rhalf$, and their halo virial radii. 
We show that in the volume-limited SDSS galaxy samples smaller galaxies cluster much more strongly than larger galaxies at fixed stellar mass. 
The size dependence increases with decreasing scale and decreases with stellar mass of the sample. Using {\tt Halotools} to forward model the observations, 
we show that a simple empirical model, in which $\rhalf$ is proptortional to the halo $\rvir$ at the time when halo mass reached its maximum,
reproduces observed size-dependence of galaxy clustering in striking detail.
The success of the model implies that the size-dependence of galaxy clustering is mainly due to the fact that satellite galaxies have smaller sizes compared 
to central galaxies of the same stellar mass. Taken together, our findings indicate that  sizes of satellite galaxies are set  prior to satellite infall and that a 
remarkably simple linear size--virial radius relation emerges from the complex physics regulating galaxy size. The $\rvir$-based size model that reproduces
 the observed size dependence of clustering makes a prediction that future observations of size dependence of galaxy-galaxy lensing for stellar mass 
 selected galaxy samples will see stronger signal for smaller galaxies. 
 %Left over this sentence. Not sure we need to talk about this model in the abstract: 
 % At the same time, model in which size is based on stellar mass does not reproduce the clustering trend, even after taking into account stellar mass loss by satellites.

\end{abstract}

@@

Observational constraints on models for galaxy size typically come from one-point function measurements such as $\mean{\rhalf}{\mstar}$ and $\sigma(\rhalf\vert\mstar),$ e.g.,  \citet{khochfar_silk06,lang_etal14,desmond_etal17,bottrell_etal17b,hou_etal17,somerville_etal17}, or otherwise from catalogs of galaxy groups \citep{weinmann_etal08,guo_etal09,huertas_company_etal13b,spindler_wake17}. More recently, the connection between a galaxy's size and the virial radius of its parent halo has been
explored by converting stellar mass to halo mass and virial radius, using abundance matching \citep{kravtsov_etal04,tasitsiomi_etal04,vale_ostriker04,vale_ostriker06,conroy_etal06}.  Such analyses have indicated that galaxy size is, on average, proportional to host halo virial radius, both at $z=0$ \citep{kravtsov13}, and higher redshifts \citep{huang_etal17,somerville_etal17}.

@@

The approach we take is to forward model the $\rhalf-$halo connection: we generate Monte Carlo realizations of models using a high-resolution cosmological $N$-body simulation, and then make synthetic measurements in a manner that closely mimics the measurements we make on our observed galaxy samples. We describe the dataset and simulations we use in \S\ref{sec:data}. Our forward-modeling methods are described in \S\ref{sec:model}, with supplementary material in the Appendix. All of our primary results appear in \S\ref{sec:results}, and their theoretical interpretation in \S\ref{sec:interpretation}. We contextualize our findings in terms of previous work in \S\ref{sec:previous_work}, and discuss future directions in \S\ref{sec:future}. We conclude the paper with a summary of our principal results in \S\ref{sec:conclusion}.

@@

As the bedrock of our modeling, we use the publicly available\footnote{\url{http://www.slac.stanford.edu/~behroozi/BPlanck\_Hlists}} catalog of {\tt Rockstar} subhalos identified at $z=0$ in the Bolshoi-Planck simulation \citep{klypin_etal11,klypin_etal16,behroozi12_rockstar,riebe_etal13,behroozi_etal12b,rodriguez_puebla16_bolplanck}. As described in \S\ref{subsec:sham}, we use traditional abundance matching to connect stellar mass $\mstar$ with masses of halos and subhalos. To address issues related to subhalo incompleteness \citep{guo_white13,campbell_etal17}, we supplement the Bolshoi-Planck subhalo catalog with subhalos that have been disrupted and no longer appear in the standard catalog as $z=0$ surviving subhalos. We describe our treatment of these ``orphan" galaxies in the Appendix, where we also show results that exclude all orphan galaxies.

For our SDSS galaxy samples, we calculate two-point clustering $\wproj$ using line-of-sight projection of $\pi_{\rm max}=20~\mpc$ using the {\tt correl} program in {\tt UniverseMachine} (Behroozi et al. 2017, in prep). The {\tt correl} code uses the \citet{landy_szalay93} estimator ($DD - 2DR + RR$) to compute the redshift-space correlation function $\xi(\rproj, \pi)$, which is then integrated over $\vert\pi\vert < 20$ Mpc to compute the projected correlation function $\wproj(\rproj)$.  The code uses $10^6$ randoms drawn from the same mask region with uniform volume distribution to compute $DR$, and $RR$ is computed via Monte Carlo integration. Errors are estimated by jackknife resampling, giving us an estimate for the lower bound of samples with volumes $V_{\rm eff}\lesssim 0.3\,{\rm Gpc}^3.$

@@

All numerical values of $\rhalf$ will be quoted in physical $\kpc,$ and all values of $\mstar$ and $\mhalo$ in $\msun,$ assuming $H_0=67.8~\kms\equiv100h~\kms,$ the best-fit value from \citet{planck15}. To scale stellar masses to ``$h=1$ units" \citep{croton13}, our numerically quoted values for $\mstar$ should be multiplied by $h^2,$ while our halo masses and distances should be multiplied by $h.$

@@

Galaxy clustering has well-known dependence upon $\mstar$, which is not the focus of this work.  We thus wish to remove this dependence to focus solely on the  $\wproj(\rproj)$ dependence
on $\rhalf.$ To do so, we determine the value $\median{\rhalf}{\mstar}$ by computing a sliding median of $\rhalf$ for galaxies sorted by $\mstar,$ calculated using a window of width $N_{\rm gal}=1000.$ \textcolor{magenta}{AK: is window symmetric? What is done for the most massive and least massive galaxies? } Each galaxy is categorized as either ``large" or ``small" according to whether it is above or below the median value appropriate for its stellar mass. We note that this is analogous to the common convention for studying the properties of ``red" vs. ``blue" galaxies, in which the two subsamples are divided by a $\mstar-$ or luminosity-dependent green valley cut \citep[e.g.,][]{vdB_etal08,zehavi_etal11}, only here the $\rhalf$ distribution is uni-modal, not bi-modal.

@@

We map $\mstar$ onto subhalos using deconvolution abundance matching (SHAM). See Section 3.3 of \citet{behroozi_etal10}, or the appendix of \citet{kravtsov_etal14}, for an account of the mathematics underlying this technique in the presence of scatter. To perform the abundance matching, we use the publicly available code\footnote{\url{https://bitbucket.org/yymao/abundancematching/overview}} developed by Yao-Yuan Mao \citep{lehmann_etal15} that provides a python wrapper around the C kernel developed in \citet{behroozi_etal10}. For the subhalo property used in SHAM, we use $\mpeak,$ the largest value of $\mvir$ ever attained by the subhalo throughout the entire history of the main progenitor halo.  Using the stellar mass function provided in \citet{moustakas_etal13}, we model $\mstar$ as a log-normal distribution with $0.2$ dex of scatter about the median relation $\median{\mstar}{\mpeak}$ determined by SHAM \citep{reddick_etal13}. As described in the appendix, we include a prescription for supplementing the ordinary {\tt Rockstar} subhalo catalog with disrupted subhalos, so that some of our model galaxies inhabit subhalos that are no longer resolved and do not appear in the standard publicly available catalog (``orphan galaxies").

@@

\subsubsection{$\mstar$-based model with satellite mass loss}
\label{subsubsec:strippingmodel}

In the first model we assume that stellar mass $\mstar$ controls $\rhalf.$ To implement this model, for simplicity we tabulate $\median{\rhalf}{\mstar}$ directly from the data, rather than pursue a parametric form, so that this model reproduces the observed $\rhalf-\mstar$ relation by construction. In order to ensure that the model has a level of scatter in $\rhalf$ at fixed $\mstar$ that is comparable to the level reported in \citet{somerville_etal17}, our model sizes are drawn from a log-normal distribution with $0.2$ dex of scatter centered at the observed $\median{\rhalf}{\mstar}.$

It is natural to consider the possibility that stellar mass is stripped from satellite galaxies after infall. The basis for our implementation of this phenomenon is the fitting function presented in \citet{smith_etal16}, which was calibrated by studying stellar mass loss in a suite of high-resolution hydrodynamical simulations. In this model, $f_{\ast}$ quantifies the remaining fraction of stellar mass as a function of $f_{\rm DM},$ the fraction of dark matter mass that remains after subhalo infall:
\beq
f_{\ast} = 1 - \exp(-14.2f_{\rm DM}).
\eeq
For $f_{\rm DM}$ we use the ratio of present-day subhalo mass divided by the peak mass, $M_{\rm vir}/M_{\rm peak}.$ If we denote the post-stripping stellar mass as $M_{\ast}',$ then we have $M_{\ast}'\equiv f_{\ast}M_{\ast},$ where $M_{\ast}$ is given by the initial application of abundance matching. We then calculate the post-stripping radius by interpolating $\langle\rhalf'\vert\mstar'\rangle$ directly from SDSS data. In this way, we model satellite mass loss and  size truncation in a manner that mimics what is seen in hydrodynamical simulations. As shown in \citet{smith_etal16}, stellar stripping in this model does not begin until $f_{\rm DM}\lesssim0.2,$ and in order for $\sim50\%$ of stars to be stripped, the DM halo needs to be stripped to $\sim5\%$ of its initial mass. 
\textcolor{magenta}{How is mass loss is handled and how stellar mass assigned for orphans? This is not described in the appendix or here.}.

@@

As noted in \S\ref{sec:intro}, \citet{kravtsov13} showed that galaxy sizes are on average proportional to the virial radius of their host halo. Motivated by these results, we  explore a model in which $\rhalf$ scales linearly with halo virial radius:\footnote{We note that the normalization factor in \citet{kravtsov13} is 0.015, whereas our normalization is 0.01. The difference derives from two factors: {\em i.} the choice of halo radius definition: \citet{bryan_norman98} in the present work, compared to the $200c$ definition used in \citet{kravtsov13}; {\em ii.} the latter study used deprojected (larger) sizes for spheroidal galaxies, whereas we make no such correction here. See \S\ref{sec:future} for further discussion.}
\beq
\label{eq:fiducial_model}
\rhalf = 0.01\rvir
\eeq
For the virial radius of halos and subhalos, we use $\rmpeak:$ the value of $\rvir$ in physical units of $\kpc$ measured at the time when halo mass reached its maximum, $\zpeak.$  The relationship between $\rmpeak, \mpeak$ and $\zpeak$ is given by
\beq
\mpeak\equiv\frac{4\pi}{3}\rmpeak^{3}\Delta_{\rm vir}(\zpeak)\rho_{\rm m}(\zpeak),
\eeq
where for $\Delta_{\rm vir}(\zpeak)$ we use the fitting function to the ``virial" definition used in \citet{bryan_norman98}. This modeling assumption effectively defines $\median{\rhalf}{\mhalo};$ this relation defines the center of the log-normal distribution with $0.2$ dex of scatter that we use to generate a Monto Carlo realization of the model population.

@@

The points and curves in Figure \ref{fig:clustering_ratio_upshot} are almost all negative: small galaxies cluster more strongly relative to large. This presentation of the measurement makes it clear that the underlying signal of $\rhalf-$dependent clustering is strongest for samples with smaller stellar mass; as $\mstar$ increases, the signal weakens and nearly vanishes for $\mstar\gtrsim10^{11}\msun.$ Strikingly, the $\rvir-$based model reproduces this $\mstar-$dependent behavior, as well as the scale-dependence of the observed clustering signal at each $\mstar.$ As described in \S\ref{subsec:censat_sizes} below, we attribute the success of this prediction to the systematic difference of sizes of central and satellite galaxies.

@@

Due to the shape of ${\rm d}\rhalf/{\rm d}\mstar,$ low-mass satellites have to lose much more halo mass than high-mass satellites in order to get pushed into the small-$\rhalf$ sample. Thus even though there are more satellites at lower masses, the shape of ${\rm d}\rhalf/{\rm d}\mstar$ counteracts this trend, making the mass-dependence of the observed signal shown in Figure \ref{fig:mstarmodelclustering} difficult to explain by appealing to tidal stripping. Evidently it is difficult to strip enough mass from satellites so that the $\rhalf-$dependent clustering predictions are in agreement with observations.  We reach the same conclusion when we tested an even more extreme $\mstar-$stripping model, in which stellar mass loss is linearly proportional to halo mass loss \citep[][Model 1]{watson_etal12}.
As we argue in \S\ref{sec:interpretation}, this supports the conclusion that the relative difference between median sizes of central and satellite galaxies is largely in place at the time of satellite infall.

@@

%---------------------------------------------------------------------------------------------------
\begin{figure}
\centering
\includegraphics[width=8cm]{FIGS/color_selected_size_dependent_clustering.pdf}
\caption{
{\bf \textcolor{magenta}{AK: it may be difficult for people (especially color-blind) to tell the difference between color of points of small and large galaxies.} Distinct $\rhalf-$dependence of clustering for color-selected galaxy samples.}
Projected clustering of SDSS galaxies in the bin of stellar mass: $10^{9.75}<\mstar/\msun<10^{10.25}.$ For both ``blue" and ``red" samples, we first divide SDSS galaxies falling in the $\mstar$ bin based on a color cut at $g-r=0.65.$ We then subsequently compute $\median{\rhalf}{\mstar; {\rm red}}$ and $\median{\rhalf}{\mstar; {\rm blue}}$. In analogy to the bottom left panel of Figure \ref{fig:sizedefinition}, for each of the red and blue samples, the small and large subsamples have statistically indistinguishable stellar mass functions. We then measure $\wproj$ for each of the four subsamples, plotting the results with color-coding as specified in the legend. At this stellar mass, pre-selecting galaxy samples by color results in a dramatic weakening of the $\rhalf-$dependence of $\wproj$ relative to $\mstar-$complete samples, as well as a sign-reversal of the signal. In \S\ref{sec:interpretation} we give a simple halo model picture unifying the trends shown here with those illustrated in Figure \ref{fig:clustering_ratio_upshot}.
}
\label{fig:colorclustering}
\end{figure}
%---------------------------------------------------------------------------------------------------

@@

Likewise, the significantly weaker dependence of clustering on size for color-selected samples (Figure \ref{fig:colorclustering}), and its reversed sign, can be naturally explained by the same principles.  It is well known that satellites are redder than centrals of the same stellar mass \citep[e.g.,][]{vdB_etal08}, and so blue galaxies have a small satellite fraction, while red galaxies have a large fraction of satellites that occupy high-mass halos; accordingly, red galaxies cluster much more strongly than blue galaxies \citep[e.g.,][]{zehavi_etal11}.

Pre-selecting galaxy samples by color then naturally restricts the impact of satellites on size-dependent clustering. For blue galaxies, the reason traces to the small value of $F_{\rm sat};$ for red galaxies, the reason traces to the limited dynamic range of host halo masses occupied by the satellites. \textcolor{magenta}{AK: I did not quite understand the explanation for red galaxies in the last sentence. It'd be good to unpack it. }. With the impact of satellites reduced, the relative clustering of centrals becomes more important. In the limit that size-dependent clustering is dominated by centrals, the $\rvir$-based size model predicts that large galaxy samples should be (weakly) more clustered than small ones, because $\rhalf\propto\rvir\propto\mhalo^{1/3}.$ This is the trend seen in the clustering measurements shown in Figure \ref{fig:colorclustering}.

@@

In the $\rvir$-based size model, the weak dependence of $\rhalf$ on $\mstar$ for low-mass galaxies is due to the steep $\mstar-\mvir$ relation. For example, if $\mstar\propto\mvir^\alpha\propto\rvir^{3\alpha}$, then $\rhalf\propto \rvir\propto \mstar^{1/(3\alpha)}$. Given that $\alpha\sim 1-3$ is inferred for the $\mstar-\mvir$ relation at low masses \citep{kravtsov10,moster_etal13,behroozi13_smhm,kravtsov_etal14}, the $\rhalf-\mstar$ relation in this regime is expected to be shallow.
For galaxies of $\mstar\gtrsim 5\times 10^{10}\ M_\odot$, $\alpha\approx 0.3$ \citep[e.g.,][]{kravtsov_etal14}, and thus
the slope of the $\rhalf-\mstar$ relation is approximately linear. These trends of $\rhalf$ with $\mstar$ have a direct relation to the trends of surface brightness $\mu_\star=\mstar/(2\pi\rhalf^2)$ with stellar mass \citep[see, for example, the $\mu_\star-$dependent clustering measurements in][which are qualitatively similar to the measurements reported here]{li_etal06}.

@@

The right panels of Figure \ref{fig:lensingprediction} compare the distributions of host halo mass, $\mhost,$ for small and large samples;\footnote{We calculated the PDFs shown in Figure \ref{fig:lensingprediction} using the Scipy implementation of Gaussian kernel density estimation \citep{scipy}.} for centrals, $\mhost=\mhalo,$ while for satellites, $\mhost$ is the virial mass of the parent halo. The peak of each distribution arises from central galaxies, while the long tail at large values of $\mhost$ is due to satellites. At all stellar masses, the blue histograms peak at slightly larger $\mhost$ relative to the red; this is because the $\rvir-$based model predicts that $\mhalo$ increases with $\rhalf$ for centrals. For $\mstar-$complete samples, this effect only becomes dominant when the satellite fraction drops to sufficiently small values. Qualitatively, these features are completely generic predictions of the $\rvir-$based model; quantitatively, the precise value of $\mstar$ where the effect of centrals begins to dominate $\Delta\Sigma$ depends sensitively on $F_{\rm sat}(\mstar),$ which we have only roughly calibrated in the present work.
